### 2020-04-14 thesis/content/chapter-related-work.tex
corrections

Before combining \ac{lta} and \ac{gcr}, as described in \cref{sec:intro:questions}, we first give an overview of the state-of-the-art in both fields of research.
This is done in three steps:
\begin{enumerate}
	\item We begin with an overview of the existing \ac{lta} methods for unstructured inputs.
	\item Then we look at the domain of structured inputs.
		To solve the \ac{gcr} problem, relevant graphs characteristics have to be defined in order to determine the similarity and dissimilarity of graphs.
		We will look at three approaches for graph characterization: The strict graph isomorphism test, the so-called \acl{wl} coloring and lastly the notion of graph spectra.
	\item Based on the described graph characterization approaches, an overview of current \ac{gcr} methods will then be given.
\end{enumerate}

@@

The class of \ac{lta} problems was first described by \citet{Melnikov2016}.
There an input instance is understood as a composition $C = \ldblbrace c_1, \dots, c_n \rdblbrace$ of so-called constituents, i.e.\ as a variable-size multiset (denoted as $\ldblbrace \cdot \rdblbrace$).
The assumption in \ac{lta} problems is that for all constituents $c_i \in C$ a local score $y_i \in \mathcal{Y}$ is either given or computable.
The set of those local scores should be indicative of the overall score $y \in \mathcal{Y}$ of the composition $C$.
\ac{lta} problems typically require two subproblems to be solved:
\begin{enumerate}[label=\textbf{\arabic*.}]
	\item \textbf{Aggregation:}
		A variadic aggregation function $\mathcal{A}: \mathcal{Y}^{*} \to \mathcal{Y}$ that estimates composition scores has to be learned, i.e.\@ $y \approx \hat{y} = \mathcal{A}(y_{1}, \dots, y_{n})$.
		Typically the aggregation function $\mathcal{A}$ should be associative and commutative to fit with the multiset-structure of compositions.
	\item \textbf{Disaggregation:}
		In case the constituent scores $y_i$ are not given, they have to be derived from a constituent representation, e.g.\ a vector $x_i \in \mathcal{X}$.
		To learn this derivation function $f: \mathcal{X} \to \mathcal{Y}$, only the constituent vectors ${\ldblbrace x_i \rdblbrace}_{i = 1}^{n}$ and the composition score $y$ is given.
		Thus the constituent scores $y_i$ need to be \textit{disaggregated} from $y$ in order to learn $f$.
\end{enumerate}
\begin{figure}[ht]
	\centering
	\includegraphics[width=0.7\linewidth]{gfx/related-work/lta-overview.pdf}
	\caption{Overview of the structure of LTA for multiset compositions.}\label{fig:related:lta-overview}
\end{figure}
Overall \ac{lta} can be understood as the joint problem of learning the aggregation function $\mathcal{A}$ and the local score derivation function $f$.
Two main approaches to represent the aggregation function in \ac{lta} problems have been explored.

@@

Since t-norms and t-conorms are commutative and associative they can also be applied to non-empty sets of arbitrary size, i.e. $T(\ldblbrace y_1, \dots, y_n \rdblbrace) = T(y_1, T(\ldblbrace y_2, \dots, y_n \rdblbrace))$ with fixpoint $T(\ldblbrace y \rdblbrace) = y$.
Using this extension, a uninorm $U_\lambda$ can be applied to sets which turns it into a parameterized aggregation function $\mathcal{A}_\lambda: {[0, 1]}^* \to [0, 1]$.
In this simple model the \ac{lta} aggregation problem boils down to the optimization of $\lambda$.
The \ac{lta} disaggregation problem is solved by jointly optimizing a \acl{lm}, i.e.\ the constituent scores ${\ldblbrace y_i \in [0, 1] \rdblbrace}_{c_i \in C}$ are described by $y_i = {\left( 1 + \exp(-\theta^\top x_i) \right)}^{-1}$.
Overall an \ac{lta} model is therefore described by the uninorm parameter $\lambda$ and the regression coefficients $\theta$.

@@

To deal with varying composition sizes $n$, the weights $\lambda_1, \dots, \lambda_n$ can however not be statically assigned.
Instead they are interpolated using a so-called \ac{bum} function $q: [0, 1] \to [0, 1]$.
It takes constituent positions that are normalized to the unit interval, i.e.\ $\frac{i}{n} \in [0, 1]$.
The \ac{bum} function $q$ is then used to interpolate a weight for any normalized sort position via $\lambda_i \coloneqq q\left(\frac{i}{n}\right) - q\left(\frac{i - 1}{n}\right)$.
Because $q$ is monotone with $q(0) = 0$ and $q(1) = 1$, it always holds that ${\|\lambda\|}_1 = q(1) - q(0) = 1$. % chktex 21
Using this model, the aggregation problem boils down to optimizing the shape of $q$.

\begin{figure}[ht]
	\centering
	\includegraphics[width=0.72\linewidth]{gfx/related-work/bum.pdf}
	\caption[Illustration of how a \ac{bum} function is described as a linear spline and its relation to the \ac{owa} weights.]{
		Illustration of how \textcolor{t_blue}{$a$} describes \textcolor{t_blue}{$q$} and its relation to \textcolor{t_red}{$\lambda$} ($n = 4$, $m = 7$).
	}\label{fig:related:bum}
\end{figure}
In the \ac{owa} approach the \ac{bum} function $q$ is modeled as a piecewise linear spline.
This spline is described by $m+1$ points ${\left\{ \left( \frac{j}{m}, a_j \right) \right\}}_{j = 0}^{m}$, the so-called knots of the spline. % chktex 21
The curve of $q$ is obtained by linearly interpolating between neighboring knots as shown in \cref{fig:related:bum}.
If $0 = a_0 \leq a_1 \leq \cdots \leq a_m = 1$, $q$ is a \ac{bum} function.
The \ac{lta} aggregation problem is therefore solved by optimizing $a \in \mathbb{R}^{m + 1}$ under this constraint.
The disaggregation problem is tackled by adding the scores $y_1, \dots, y_M \in \mathbb{R}$ to the learnable parameters of the model where $M$ is assumed to be the finite number of constituents.
Currently the \ac{owa} approach requires all possible constituents to be part of the training dataset since it does not consider constituent features $x_i \in \mathcal{X}$ to predict the scores of previously unseen constituents.

@@

\begin{defn}\label{defn:related:graph}
	A \textit{graph} $G \coloneqq (\mathcal{V}_G, \mathcal{E}_G)$ consists of a finite set of vertices $v_i \in \mathcal{V}_G$ and a set of edges $e_{ij} = (v_i, v_j) \in \mathcal{E}_G \subseteq \mathcal{V}_G^2$.
	Optionally discrete vertex labels $l_G[v_i] \in L_{\mathcal{V}}$ or edge labels $l_G[e_{ij}] \in L_{\mathcal{E}}$ may be associated with all vertices $v_i \in \mathcal{V}_G$ and edges $e_{ij}\in \mathcal{E}_G$ respectively.
	Also continuous feature vectors $x_G[v_{i}] \in \mathcal{X}_{\mathcal{V}}$, $x_G[e_{ij}] \in \mathcal{X}_{\mathcal{E}}$ may be given.
	If $\mathcal{X}_{\mathcal{E}} = \mathbb{R}$, $x_G[e_{ij}]$ can be interpreted as an edge weight of $e_{ij}$.

	In this thesis all graphs $G$ are assumed to be undirected if not explicitly stated otherwise, i.e.\ ${e_{ij} \in \mathcal{E}_G} \leftrightarrow\allowbreak {e_{ji} \in \mathcal{E}_G} \allowbreak\land l_G[e_{ij}] = l_G[e_{ji}] \allowbreak\land x_G[e_{ij}] = x_G[e_{ji}]$.
	We denote the set of all undirected graphs as $\mathcal{G}$.
	Additionally we denote $G[S] \coloneqq (S, \mathcal{E}_G \cap S^2)$ as the subgraph of $G$ induced by $S \subseteq \mathcal{V}_G$.
\end{defn}
To classify or score a graph, it first needs to be characterized by a set of relevant properties.
The most strict characterization of a graph is its so-called \textit{isomorphism class}.
It uniquely identifies a graph but lacks any notion of similarity between non-identical graphs.
We will begin with a brief definition of this strict isomorphism-based graph characterization.
Then two less strict characterization approaches are described; the so called \acl{wl} coloring and the notion of graph spectra.
They are the theoretical foundation of many current \ac{gcr} methods.

@@

In order to process a graph $G$ of size $n = |\mathcal{V}_G|$, one is forced to choose some encoding, e.g.\ an adjacency matrix $A_G \in {\{0, 1\}}^{n \times n}$.
Such an encoding introduces a vertex ordering $v_1, \dots, v_n$ that does not carry any semantic meaning.
Consequently there are $n!$ equivalent encodings of $G$. % chktex 40
To represent those encodings we introduce the notion of ordered induced subgraphs.
\begin{defn}\label{defn:related:ordered-subgraph}
	For all vertex $k$-tuples $v = (v_1, \dots, v_k) \in \mathcal{V}_G^k$ let $\hat{v} = \set(v) \coloneqq \{ v_i\, |\, i \in [k] \}$.
	Then $G[v] \coloneqq (\hat{v}, \mathcal{E}_G \cap \hat{v}^2, v)$ is called the \textit{ordered subgraph} of $G$ induced by $v$.
	We denote the set of all ordered subgraphs with $\mathcal{G}_{\mathit{ord}}$.
	If $\hat{v} = \mathcal{V}_G$ and $k = |\mathcal{V}_G|$, we call $v$ an \textit{ordering} of $G$ and $G[v]$ an \textit{encoding} of $G$.
\end{defn}
\begin{defn}\label{defn:related:ordered-subgraph-equivalence}
	Two ordered induced subgraphs $G[v]$ and $H[w]$ with $(v_1, \dots, v_k) \in \mathcal{V}_G^k$ and $(w_1, \dots, w_k) \in \mathcal{V}_H^k$ are \textit{equivalent} ($G[v] \equiv H[w]$) iff.\
	\begin{align*}
		&\quad\forall i, j \in [k]: (v_i, v_j) \in \mathcal{E}_G \leftrightarrow (w_i, w_j) \in \mathcal{E}_H \\
		&\, \begin{aligned}
			\land&\ \forall i, j \in [k]: l_G[v_i, v_j] = l_H[w_i, w_j] &\land&\quad\forall i \in [k]: l_G[v_i] = l_H[w_i] \\
			\land&\ \forall i, j \in [k]: x_G[v_i, v_j] = x_H[w_i, w_j] &\land&\quad\forall i \in [k]: x_G[v_i] = x_H[w_i]\text{.}
		\end{aligned}
	\end{align*}
	Using this notion of equivalence, we call ${\left[ G[v] \right]} \coloneqq \{ G[w]\, |\, G[w] \equiv G[v] \land w \in \mathcal{V}_G^{*} \}$ the \textit{ordered subgraph equivalence class} of $G[v]$.
\end{defn}
\begin{defn}
	Two graphs $G$ and $H$ are \textit{isomorphic} ($G \simeq H$) iff.\ there are equivalent encodings $G[v] \equiv H[w]$ of them.
	Consequently $[G] \coloneqq \{ H\, |\, H \simeq G \}$ is called the \textit{isomorphism class} of $G$.
\end{defn}
The goal of the \ac{gi} problem is to check whether $G \simeq H$ for two arbitrary graphs.
Even though there is no known universal polynomial algorithm that solves \ac{gi}, \citet{Babai1980} showed that almost all graphs can be trivially distinguished in linear time.
More recently \citet{Babai2015} also presented a quasipolynomial upper time bound for the remaining hard \ac{gi} instances.
For all practical purposes, \ac{gi} can therefore be solved efficiently; e.g.\ via the \texttt{nauty} program~\cite{McKay2013}\cite{McKay}.
One important subroutine in most \ac{gi} checkers is the \acl{wl} algorithm which will be described next.

@@

In addition to the binary aspect of \ac{wl} distinguishability and its relation to the \ac{gi} problem, \ac{wl} colorings are also useful for more fuzzy graph similarity comparisons as we will see in \cref{sec:related:gcr:kernel} when we look at graph kernels.
Before that however, the details of \ac{wl} color refinement strategy have to be described.
We begin with the color refinement algorithm for the most simple case of $k = 1$.
Then the definitions and intuitions from the 1-dimensional case are extended to its higher-dimensional generalization.
Lastly we will discuss the discriminative power of the \acs{wl} algorithm and its relation to the \acs{wl}-dimension $k$.

\subsubsection{The 1-dimensional \acs{wl} algorithm}
In the 1-dimensional \ac{wl} algorithm (1-WL), a color is assigned to each vertex of a graph.
If the vertices $v \in \mathcal{V}_G$ of the input graph $G$ are labeled, those labels $l_G[v] \in L_{\mathcal{V}} \subseteq \mathcal{C}$ can be used as the initial graph coloring $\chi_{G,1}^{(0)}(v) \coloneqq l_G[v]$.
Since \ac{wl} colors are inherently discrete, continuous vertex feature vectors $x_G[v]$ are not considered here.
For unlabeled graphs a constant coloring is used, e.g.\ $\forall v \in \mathcal{V}_G: \chi_{G,1}^{(0)}(v) = \colorlabel{t_blue}{A}$ for some initial color $\colorlabel{t_blue}{A} \in \mathcal{C}$. % chktex 25
In each iteration of the 1-\acs{wl} color refinement algorithm, the following neighborhood aggregation scheme is used to compute a new color for all vertices:
\begin{defn}\label{defn:related:wl1-refine}
	$\chi_{G,1}^{(i+1)}(v) \coloneqq h\left(\chi_{G,1}^{(i)}(v), \ldblbrace \chi_{G,1}^{(i)}(u)\, |\, u \in \Gamma_{G}(v) \rdblbrace\right)$,
	with $\Gamma_G(v)$ denoting the set of adjacent vertices of $v \in \mathcal{V}_G$ and $h: \mathcal{C}^* \to \mathcal{C}$ denoting an injective hash function that assigns a unique color to each finite combination of colors.
\end{defn}
\begin{figure}[ht]
	\centering
	\includegraphics[width=0.72\linewidth]{gfx/related-work/wl1-refine.pdf}
	\caption[Example 1-WL color refinement steps.]{
		Example 1-WL color refinement steps.
		After two iterations the coloring stabilizes.
		Each vertex $v$ is labeled with its current color and has its previous color and the colors of the hashed neighbors $\Gamma_G(v)$ written next to it (see \cref{defn:related:wl1-refine}).
	}\label{fig:related:wl1-refine}
\end{figure}
In practice the hash function $h$ is usually defined lazily by using $\mathcal{C} \subseteq \mathbb{N}$ and enumerating color combinations in whichever order they are hashed s.t.\ a new color is introduced every time a previously unseen color combination appears at runtime\footnote{
	$\mathcal{C}$ is countably infinite but we assume it only contains the finite no.\ of colors occurring in a dataset.
}.

>>>>

Before combining \ac{lta} and \ac{gcr} as described in \cref{sec:intro:questions}, we first give an overview of the state-of-the-art in both fields of research.
This is done in three steps:
\begin{enumerate}
	\item We begin with an overview of the existing \ac{lta} methods for unstructured inputs.
	\item Then we look at the domain of structured inputs.
		To solve the \ac{gcr} problem, relevant graphs characteristics have to be defined in order to determine the similarity and dissimilarity of graphs.
		We will look at three approaches for graph characterization: The strict graph isomorphism test, the so-called \acl{wl} coloring and lastly the notion of graph spectra.
	\item Based on the described graph characterization approaches, an overview of current \ac{gcr} methods will then be given.
\end{enumerate}

@@

The class of \ac{lta} problems was first described by \citet{Melnikov2016}.
There an input instance is understood as a composition $C = \ldblbrace c_1, \dots, c_n \rdblbrace$ of so-called constituents, i.e.\ as a variable-size multiset (denoted as $\ldblbrace \cdot \rdblbrace$).
The assumption in \ac{lta} problems is that, for all constituents $c_i \in C$, a local score $y_i \in \mathcal{Y}$ is either given or computable.
The set of those local scores should be indicative of the overall score $y \in \mathcal{Y}$ of the composition $C$.
\ac{lta} problems typically require two subproblems to be solved:
\begin{enumerate}[label=\textbf{\arabic*.}]
	\item \textbf{Aggregation:}
		A variadic aggregation function $\mathcal{A}: \mathcal{Y}^{*} \to \mathcal{Y}$ that estimates composition scores has to be learned, i.e.\@ $y \approx \hat{y} = \mathcal{A}(y_{1}, \dots, y_{n})$.
		Typically the aggregation function $\mathcal{A}$ should be associative and commutative to fit with the multiset-structure of compositions.
	\item \textbf{Disaggregation:}
		In case the constituent scores $y_i$ are not given, they have to be derived from a constituent representation, e.g.\ a vector $x_i \in \mathcal{X}$.
		To learn this derivation function $f: \mathcal{X} \to \mathcal{Y}$, only the constituent vectors ${\ldblbrace x_i \rdblbrace}_{i = 1}^{n}$ and the composition score $y$ is given.
		Thus the constituent scores $y_i$ need to be \textit{disaggregated} from $y$ in order to learn $f$.
\end{enumerate}
\begin{figure}[ht]
	\centering
	\includegraphics[width=0.7\linewidth]{gfx/related-work/lta-overview.pdf}
	\caption{Overview of the structure of LTA for multiset compositions.}\label{fig:related:lta-overview}
\end{figure}
Overall, \ac{lta} can be understood as the joint problem of learning the aggregation function $\mathcal{A}$ and the local score derivation function $f$.
Two main approaches to represent the aggregation function in \ac{lta} problems have been explored.

@@

Since t-norms and t-conorms are commutative and associative, they can also be applied to non-empty sets of arbitrary size, i.e. $T(\ldblbrace y_1, \dots, y_n \rdblbrace) = T(y_1, T(\ldblbrace y_2, \dots, y_n \rdblbrace))$ with fixpoint $T(\ldblbrace y \rdblbrace) = y$.
Using this extension, a uninorm $U_\lambda$ can be applied to sets, which turns it into a parameterized aggregation function $\mathcal{A}_\lambda: {[0, 1]}^* \to [0, 1]$.
In this simple model the \ac{lta} aggregation problem boils down to the optimization of $\lambda$.
The \ac{lta} disaggregation problem is solved by jointly optimizing a \acl{lm}; i.e.\ the constituent scores ${\ldblbrace y_i \in [0, 1] \rdblbrace}_{c_i \in C}$ are described by $y_i = {\left( 1 + \exp(-\theta^\top x_i) \right)}^{-1}$.
Overall, such an \ac{lta} model is therefore described by the uninorm parameter $\lambda$ and the regression coefficients $\theta$.

@@

To deal with varying composition sizes $n$, the weights $\lambda_1, \dots, \lambda_n$ can however not be statically assigned.
Instead, they are interpolated using a so-called \ac{bum} function $q: [0, 1] \to [0, 1]$ which takes constituent positions that are normalized to the unit interval, i.e.\ $\frac{i}{n} \in [0, 1]$.
The \ac{bum} function $q$ is used to interpolate a weight for any normalized sort position via $\lambda_i \coloneqq q\left(\frac{i}{n}\right) - q\left(\frac{i - 1}{n}\right)$.
Because $q$ is monotone with $q(0) = 0$ and $q(1) = 1$, it always holds that ${\|\lambda\|}_1 = q(1) - q(0) = 1$. % chktex 21
Using this model, the aggregation problem boils down to optimizing the shape of $q$.

\begin{figure}[ht]
	\centering
	\includegraphics[width=0.72\linewidth]{gfx/related-work/bum.pdf}
	\caption[Illustration of how a \ac{bum} function is described as a linear spline and its relation to the \ac{owa} weights.]{
		Illustration of how \textcolor{t_blue}{$a$} describes \textcolor{t_blue}{$q$} and its relation to \textcolor{t_red}{$\lambda$} ($n = 4$, $m = 7$).
	}\label{fig:related:bum}
\end{figure}
In the \ac{owa} \ac{lta} approach the \ac{bum} function $q$ is modeled as a piecewise linear spline.
This spline is described by $m+1$ points ${\left\{ \left( \frac{j}{m}, a_j \right) \right\}}_{j = 0}^{m}$, the so-called knots of the spline. % chktex 21
The curve of $q$ is obtained by linearly interpolating between neighboring knots as shown in \cref{fig:related:bum}.
If $0 = a_0 \leq a_1 \leq \cdots \leq a_m = 1$, $q$ is a \ac{bum} function.
The \ac{lta} aggregation problem is therefore solved by optimizing $a \in \mathbb{R}^{m + 1}$ under this constraint.
The disaggregation problem is tackled by adding the scores $y_1, \dots, y_M \in \mathbb{R}$ to the learnable parameters of the model, where $M$ is assumed to be the finite number of constituents.
Currently the \ac{owa} approach requires all possible constituents to be part of the training dataset since it does not consider constituent features $x_i \in \mathcal{X}$ to predict the scores of previously unseen constituents.

@@

\begin{defn}\label{defn:related:graph}
	A \textit{graph} $G \coloneqq (\mathcal{V}_G, \mathcal{E}_G)$ consists of a finite set of vertices $v_i \in \mathcal{V}_G$ and a set of edges $e_{ij} = (v_i, v_j) \in \mathcal{E}_G \subseteq \mathcal{V}_G^2$.
	Optionally, discrete vertex labels $l_G[v_i] \in L_{\mathcal{V}}$ or edge labels $l_G[e_{ij}] \in L_{\mathcal{E}}$ may be associated with all vertices $v_i \in \mathcal{V}_G$ and edges $e_{ij}\in \mathcal{E}_G$ respectively.
	Also continuous feature vectors $x_G[v_{i}] \in \mathcal{X}_{\mathcal{V}}$, $x_G[e_{ij}] \in \mathcal{X}_{\mathcal{E}}$ may be given.
	If $\mathcal{X}_{\mathcal{E}} = \mathbb{R}$, $x_G[e_{ij}]$ can be interpreted as an edge weight of $e_{ij}$.

	In this thesis, all graphs $G$ are assumed to be undirected if not explicitly stated otherwise, i.e.\ ${e_{ij} \in \mathcal{E}_G} \leftrightarrow\allowbreak {e_{ji} \in \mathcal{E}_G} \allowbreak\land l_G[e_{ij}] = l_G[e_{ji}] \allowbreak\land x_G[e_{ij}] = x_G[e_{ji}]$.
	We denote the set of all undirected graphs as $\mathcal{G}$.
	Additionally, we denote $G[S] \coloneqq (S, \mathcal{E}_G \cap S^2)$ as the subgraph of $G$ induced by $S \subseteq \mathcal{V}_G$.
\end{defn}
To classify or score a graph, it first needs to be characterized by a set of relevant properties.
The most strict characterization of a graph is its so-called \textit{isomorphism class}.
It uniquely identifies a graph but lacks any notion of similarity between non-identical graphs.
We will begin with a brief definition of this strict isomorphism-based graph characterization.
Then two less strict characterization approaches are described; the so called \acl{wl} coloring and the notion of graph spectra.
They are the theoretical foundation of many current \ac{gcr} methods.

@@

In order to process a graph $G$ of size $n = |\mathcal{V}_G|$, one is forced to choose some encoding, e.g.\ an adjacency matrix $A_G \in {\{0, 1\}}^{n \times n}$.
Such an encoding introduces a vertex ordering $v_1, \dots, v_n$ that does not carry any semantic meaning.
Consequently there are $n!$ equivalent encodings of $G$. % chktex 40
To represent those encodings, we introduce the notion of ordered induced subgraphs.
\begin{defn}\label{defn:related:ordered-subgraph}
	For all vertex $k$-tuples $v = (v_1, \dots, v_k) \in \mathcal{V}_G^k$, let $\hat{v} = \set(v) \coloneqq \{ v_i\, |\, i \in [k] \}$.
	Then $G[v] \coloneqq (\hat{v}, \mathcal{E}_G \cap \hat{v}^2, v)$ is called the \textit{ordered subgraph} of $G$ induced by $v$.
	We denote the set of all ordered subgraphs with $\mathcal{G}_{\mathit{ord}}$.
	If $\hat{v} = \mathcal{V}_G$ and $k = |\mathcal{V}_G|$, we call $v$ an \textit{ordering} of $G$ and $G[v]$ an \textit{encoding} of $G$.
\end{defn}
\begin{defn}\label{defn:related:ordered-subgraph-equivalence}
	Two ordered induced subgraphs $G[v]$ and $H[w]$ with $(v_1, \dots, v_k) \in \mathcal{V}_G^k$ and $(w_1, \dots, w_k) \in \mathcal{V}_H^k$ are \textit{equivalent} ($G[v] \equiv H[w]$) iff.\
	\begin{align*}
		&\quad\forall i, j \in [k]: (v_i, v_j) \in \mathcal{E}_G \leftrightarrow (w_i, w_j) \in \mathcal{E}_H \\
		&\, \begin{aligned}
			\land&\ \forall i, j \in [k]: l_G[v_i, v_j] = l_H[w_i, w_j] &\land&\quad\forall i \in [k]: l_G[v_i] = l_H[w_i] \\
			\land&\ \forall i, j \in [k]: x_G[v_i, v_j] = x_H[w_i, w_j] &\land&\quad\forall i \in [k]: x_G[v_i] = x_H[w_i]\text{.}
		\end{aligned}
	\end{align*}
	Using this notion of equivalence, we call ${\left[ G[v] \right]} \coloneqq \{ G[w]\, |\, G[w] \equiv G[v] \land w \in \mathcal{V}_G^{*} \}$ the \textit{ordered subgraph equivalence class} of $G[v]$.
\end{defn}
\begin{defn}
	Two graphs $G$ and $H$ are \textit{isomorphic} ($G \simeq H$) iff.\ there are equivalent encodings $G[v] \equiv H[w]$ of them.
	Consequently $[G] \coloneqq \{ H\, |\, H \simeq G \}$ is called the \textit{isomorphism class} of $G$.
\end{defn}
The goal of the \ac{gi} problem is to check whether $G \simeq H$ for two arbitrary graphs.
Even though there is no known universal polynomial algorithm that solves \ac{gi}, \citet{Babai1980} showed that almost all graphs can be trivially distinguished in linear time.
More recently \citet{Babai2015} also presented a quasipolynomial upper time bound for the remaining hard \ac{gi} instances.
For all practical purposes, \ac{gi} can therefore be solved efficiently; e.g.\ via the \texttt{nauty} program~\cite{McKay2013}\cite{McKay}.
One important subroutine in most \ac{gi} checkers is the \acl{wl} algorithm which will be described next.

@@

In addition to the binary aspect of \ac{wl} distinguishability and its relation to the \ac{gi} problem, \ac{wl} colorings are also useful for more fuzzy graph similarity comparisons as we will see in \cref{sec:related:gcr:kernel} when we look at graph kernels.
Before that however, the details of the \ac{wl} color refinement strategy have to be described.
We begin with the color refinement algorithm for the most simple case of $k = 1$.
Then the definitions and intuitions from the 1-dimensional case are extended to its higher-dimensional generalization.
Lastly we will discuss the discriminative power of the \acs{wl} algorithm and its relation to the \acs{wl}-dimension $k$.

\subsubsection{The 1-dimensional \acs{wl} algorithm}
In the 1-dimensional \ac{wl} algorithm (1-WL), a color is assigned to each vertex of a graph.
If the vertices $v \in \mathcal{V}_G$ of the input graph $G$ are labeled, those labels $l_G[v] \in L_{\mathcal{V}} \subseteq \mathcal{C}$ can be used as the initial graph coloring $\chi_{G,1}^{(0)}(v) \coloneqq l_G[v]$.
Since \ac{wl} colors are inherently discrete, continuous vertex feature vectors $x_G[v]$ are not considered here.
For unlabeled graphs a constant coloring is used, e.g.\ $\forall v \in \mathcal{V}_G: \chi_{G,1}^{(0)}(v) = \colorlabel{t_blue}{A}$ for some initial color $\colorlabel{t_blue}{A} \in \mathcal{C}$. % chktex 25
In each iteration of the 1-\acs{wl} color refinement algorithm, the following neighborhood aggregation scheme is used to compute a new color for each vertex:
\begin{defn}\label{defn:related:wl1-refine}
	$\chi_{G,1}^{(i+1)}(v) \coloneqq h\left(\chi_{G,1}^{(i)}(v), \ldblbrace \chi_{G,1}^{(i)}(u)\, |\, u \in \Gamma_{G}(v) \rdblbrace\right)$,
	with $\Gamma_G(v)$ denoting the set of adjacent vertices of $v \in \mathcal{V}_G$ and $h: \mathcal{C}^* \to \mathcal{C}$ denoting an injective hash function that assigns a unique color to each finite combination of colors.
\end{defn}
\begin{figure}[ht]
	\centering
	\includegraphics[width=0.72\linewidth]{gfx/related-work/wl1-refine.pdf}
	\caption[Example 1-WL color refinement steps.]{
		Example 1-WL color refinement steps.
		After two iterations the coloring stabilizes.
		Each vertex $v$ is labeled with its current color and has its previous color and the colors of the hashed neighbors $\Gamma_G(v)$ written next to it (see \cref{defn:related:wl1-refine}).
	}\label{fig:related:wl1-refine}
\end{figure}
In practice the hash function $h$ is usually defined lazily by using $\mathcal{C} \subseteq \mathbb{N}$ and enumerating color combinations in whichever order they are hashed s.t.\ a new color is introduced every time a previously unseen color combination appears at runtime\footnote{
	$\mathcal{C}$ is countably infinite but we assume it only contains the finite no.\ of colors occurring in a dataset.
}.

\subsubsection{The $k$-dimensional \acs{wl} algorithm}
As we just saw, the 1-\acs{wl} algorithm iteratively refines colorings of single vertices.
While the obtained colorings differ for most non-isomorphic graphs $G \not\simeq H$, 1-\acs{wl} does not generally solve the \ac{gi} problem as illustrated in \cref{fig:related:wl1-problem}.
\begin{figure}[ht]
	\centering
	\includegraphics[width=\linewidth]{gfx/related-work/wl1-problem.pdf}
	\caption[Two simple non-isomorphic graphs that are indistinguishable by 1-\acs{wl}.]{
		Two simple non-isomorphic graphs that are indistinguishable by 1-\acs{wl}.
	}\label{fig:related:wl1-problem}
\end{figure}\\
By extending \ac{wl} to higher dimensions, such 1-\acs{wl} indistinguishable cases can however be handled.
Analogous to the 1-dimensional variant in \cref{defn:related:wl1-refine}, the $k$-dimensional color refinement step is defined by
\begin{defn}\label{defn:related:wlk-refine}
	$\begin{aligned}[t]
		\chi_{G,k}^{(i+1)}(s) \coloneqq h\left(\chi_{G,k}^{(i)}(s), \ldblbrace (\chi_{G,k}^{(i)}(s[u/1]), \dots, \chi_{G,k}^{(i)}(s[u/k]))\, |\, u \in \mathcal{V}_G \rdblbrace\right)\\
		\text{with } s = (v_1, \dots, v_k) \in \mathcal{V}_G^k,\quad s[u/j] \coloneqq (v_1, \dots, v_{j-1}, u, v_{j+1}, \dots, v_k) \text{.}
	\end{aligned}$
\end{defn}
In 1-\acs{wl}, a vertex color is refined by combining the colors of neighboring vertices.
In $k$-\acs{wl}, the color of a $k$-tuple $s \in \mathcal{V}_G^k$ is refined by combining the colors of its neighborhood which is defined as the set of all $k$-tuples in which at most one vertex differs from $s$.
Note that each vertex $k$-tuple has one neighbor for each $u \in \mathcal{V}_G$, each of which is a $k$-tuple of vertex $k$-tuples.
This more abstract notion of neighborhood is illustrated in \cref{fig:related:wl-neighbors}.
\begin{figure}[ht]
	\centering
	\begin{subfigure}{0.32\textwidth}
		\centering
		\includegraphics[width=0.8\textwidth]{gfx/related-work/wl1-neighbors.pdf}
		\subcaption{$v_1$'s 1-WL neighbors}\label{fig:related:wl-neighbors:1}
	\end{subfigure}%
	\begin{subfigure}{0.32\textwidth}
		\centering
		\includegraphics[width=0.8\textwidth]{gfx/related-work/wl2-neighbors.pdf}
		\subcaption{$(v_1, v_2)$'s 2-WL neighbors}\label{fig:related:wl-neighbors:2}
	\end{subfigure}%
	\begin{subfigure}{0.36\textwidth}
		\centering
		\includegraphics[width=0.711\textwidth]{gfx/related-work/wl3-neighbors.pdf}
		\subcaption{$(v_1, v_2, v_3)$'s 3-WL neighbors}\label{fig:related:wl-neighbors:3}
	\end{subfigure}
	\caption[\ac{wl} neighborhoods for different values of $k$.]{
		Tuple neighborhoods for different values of $k$.
		The vertices highlighted in \textcolor{t_blue}{blue} form the root tuple $s \in \mathcal{V}_G^k$ whose neighbors are shown;
		for simplicity neighbors with $u \in s$ are left out (see~\cref{defn:related:wlk-refine}).
		Each neighbor is highlighted with a different color, except for 3-WL where the \textcolor{t_red}{red}, \textcolor{t_darkgreen}{green} and \textcolor{t_darkyellow}{yellow} triples actually form the single neighbor for $u = v_4$.
	}\label{fig:related:wl-neighbors}
\end{figure}
For $k = 2$ this means that each potential edge $(v, w) \in \mathcal{V}_G^2$ has all possible walks of length $2$ from $v$ to $w$ as its neighbors (see~\cref{fig:related:wl-neighbors:2}).
Also note that, even though $k$-\acs{wl} refines $k$-tuple colors, lower-dimensional structures still get their own colors since a tuple does not have to consist of distinct vertices, i.e.\ in $k$-\acs{wl} the color of a single vertex $v \in \mathcal{V}_G$ is described by $\chi^{*}_{G, k}(s)$ for $s = (v, \dots, v) \in \mathcal{V}_G^k$.

Let us now look at how the tuple colors are initialized.
For this we use the ordered subgraph equivalence classes $[G[s]]$ (see~\cref{defn:related:ordered-subgraph-equivalence}) which determine the initial color $\chi_{G,k}^{(0)}(v)$ of each $k$-tuple $s$.
For $k = 1$ the equivalence class of a single vertex $v$ directly corresponds to its label $l_G[v]$.
More generally, for $k > 1$ this means that
\begin{align}
	\chi_{G,k}^{(0)}(s) = \chi_{G,k}^{(0)}(t) \iff G[s] \equiv G[t] \text{.} \label{eq:related:wlk-init}
\end{align}
Note that there is a fundamental difference in how the adjacency information encoded in $\mathcal{E}_G$ is used in 1-\acs{wl} vs.\ $k$-\acs{wl}:
In 1-\acs{wl} a vertex coloring by itself cannot encode adjacency which is why this information is explicitly incorporated in each refinement step via $\Gamma_G$ (see~\cref{defn:related:wl1-refine}).
In $k$-\acs{wl} on the other hand, each pair of vertices $(v, u) \in \mathcal{V}_G^2$ appears in at least one $k$-tuple (assuming $k \geq 2$) and therefore has at least one color which can implicitly encode the adjacency information.
Edges and non-edges are colored differently in the initial coloring since $G[(v, u)] \not\equiv G[(w, u)]$ if $(v, u) \in \mathcal{E}_G$ but $(w, u) \notin \mathcal{E}_G$;
thus no explicit adjacency information is needed in the $k$-\acs{wl} color refinement step in \cref{defn:related:wlk-refine}.
